---
title: "Data Preparation is Imporatant: An Example Using SATSA"
author: "Jamie Knight"
date: "University of Victoria"
output: 
  ioslides_presentation:
  fig_width: 7
  fig_height: 6
incremental: "true" 
logo: jamieknighticon.jpg
---


```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_preparation/data_preparationFINAL.R") # the file to which knitr calls for the chunks
```


##Introduction {.build}

Hello!

Things we will cover today:  

- Why is data preparation imporatant
- An example
- Introduction to SATSA data set


##Why is data preparation imporatant? {.build}

* You will spend the majority of your time prepping your data set
  + Do this well to get good, reproducible results
  + Any errors are likely going to affect your analysis
* Some of you will use other people's data: Don't assume it is clean!
  + Even well used and documented data may have problems


##An Example 

###Let's examine an example of why data preparation is imporatant using the SATSA data.###

SATSA is freely available on the internet.
It is well used and there are over 30,000 publications using this data set!
You can access it by signing up on their website here:

[http://www.icpsr.umich.edu/icpsrweb/DSDR/studies/3843](http://www.icpsr.umich.edu/icpsrweb/DSDR/studies/3843)


##Introduction to SATSA
The data set is comprised of twins raised together and apart and includes data on:  

- Health Status
- Work Environment
- Personality

##Data for this Project {.build}

A subsample of the participants were interviewed in seven waves of in-person testing and those measurements included smell and taste questions along with cognitive and functional capacity.

This subsample is what we will use in our analysis and includes:

>- Questionnaire 3 and In-Person Testing Time 3 (1990); 
>- In-Person Testing Time 5 (2004); 
>- Questionnaire 6 and In-Person Testing Time 6 (2007); 
>- In-Person Testing Time 7 (2010)

##My Research

Can olfactory decline predict neurodegenerative disease?

A good use for SATSA

##Data Preparation

Each wave of SATSA is in a separate data set and combining several waves of data requires a substantial amount of work.

> In the folder called reports we have a script dedicated to how we merged 4 waves of in person data colected from questionnaires with the administrative and cognitive data sets.

[reports/data_preparation/data_preparationFINAL.Rmd](reports/data_preparation/data_preparationFINAL.Rmd)


##Here comes the problem

At the point where I merge the data you can start to see where the error is.

The variables have been coded using different responses over each of the waves.
This makes longitudinal analysis impossible for these variables.

```{r echo=FALSE, message=C_message, warning=C_warning}
pathDir <- getwd()
load(file.path(pathDir,"data/derived/dslong.rda"))
load(file.path(pathDir,"data/derived/idfsample2.rda"))
```

```{r show_values, echo=FALSE, message=C_message, warning=C_warning}
```


##Can You Spot the Problem?

```{r LoadPackages, echo=FALSE, message=C_message, warning=C_warning}
```

```{r echo=FALSE, message=C_message, warning=C_warning}
pathDir <- getwd()
load(file.path(pathDir,"data/derived/dslong.rda"))
```

```{r show_values, echo=FALSE, fig.show='animate', message=C_message, warning=C_warning}
```

##How to deal with issues like this {.build}
>1. Contact your supervisor 
>2. Contact the PI for the data set
>3. Fix the problem (for everyone not just your publication)

##Data Analsyis
**My Research**

I am interested in whether examining whether we can use a decline in smell to predict dementia.

Using the SATSA data I explore the change in smell over time. I am currently missing the diagnoses data (this will be my outcome variable).

##Visual | Exploring how a person changes over time
Fitted OLS trajectories superimposed on empirical growth plots
```{r echo=FALSE, message=C_message, warning=C_warning}
xyplot(smell ~ WAVE | TWINNR, data=idfsample2, 
       panel = function(x, y){
         panel.xyplot(x, y)
         panel.lmline(x, y)
       }, ylim=c(0, 5), as.table=T)
```

##next



