---
title: "Dial Post Test- Data Preparation"
output:
  html_document:
    toc: yes
    self_containted: false
    keep_md: true
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: the-name-of-the-repository
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_creation/dialr.R") # the file to which knitr calls for the chunks
```

```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
table(ds)
```

```{r DeclareGlobals, echo=C_echo, message=C_message, warning=C_warning}
```
Load tagged data from Superlab
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

Rename tagged data to be more workable in R while remaning sufficiently descriptive 
```{r RenameVariables, echo=C_echo, message=C_message, warning=C_warning}
```
Inspection of reaction time data to find outliers that are likely not refelctive of real responses. 
```{r QueryData, echo=C_echo, message=C_message, warning=C_warning}
```
Apply a filter to remove trials with a total RT<1000ms (these are behaviourally speaking not possible and must be due to an accidental response before task completion) as well as reaction times greater than 20000ms as these are upper-bound outliers and reflect less than 0.005% of the total dataset. 
```{r Filter, echo=C_echo, message=C_message, warning=C_warning}
```
Aggregate data such that for each subject a mean per-charatcer-rt by for each sentence type is reported, as well as an accuracy rating by sentence type
```{r Aggregate1, echo=C_echo, message=C_message, warning=C_warning}
```
Aggregate data such that for each sentence a mean per-charatcer-rt by for each sentence type is reported, as well as an accuracy rating by sentence type. Note that sentence here refers not to the precise sentence stimuli presented, but the temporal position of the sentence within the experiment. 
```{r Aggregate2, echo=C_echo, message=C_message, warning=C_warning}
```

Save derived data so prepare for data analysis and graphing (in separate report)
```{r SaveDerivedData, echo=C_echo, message=C_message, warning=C_warning}
```
Clean up workspace 
```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```

