---
title: "Final Project"
author: "Thomas Ferguson"
date: "December 3, 2015"
output: 
  ioslides_presentation: 
    smaller: yes
    widescreen: yes
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- FALSE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_analysis/basic_analysis_segmented.R") # the file to which knitr calls for the chunks
```

##About the project

This project attempts to model BMI (Body mass index) versus a number of different predictor variables (such as educatior (in years), age, cognition score (a measure that combines working memory and a number of other variables)). Best sub-set selection is used to identify the ideal number of variables to include in the model. Cross-validation was attempted but did not work due to technical errors. Instead, 4 coefficiants (Rss, Rsquared, BIC and Cp) were calculated to assess model fit. Significant variables had their shape (linear, quadratic, cubic, etc.) assessed to determine the most appropriate relationship to BMI. In the end, no multivariate models were modeled, however a number of univariate models of variables that were identified to be significant were used (using the appropriate variable form determined above) and had their relationship to BMI plotted (along with factors that were determined to be significant such as if the participant drank, race or gender)

##Load Packages
Prepare RStudio environment for all tasks to follow. 
```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```

```{r load_sources, echo=C_echo, message=C_message, warning=C_warning}
```

```{r DeclareGlobals, echo=C_echo, message=C_message, warning=C_warning}
```

We load the data derived by the script "./reports/data_preparation/dsL_hrs.R"" 
Variables chosen are age in years, number of health conditions, eduction in years and self-rated health score
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 1

Load graph settings for creating figures
```{r LoadGraphSettings, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 2
Chooses variables for model - End up with model of ~6 or 7 due to conflicting results (seemed reasonable - dropped the least significant factors which were gender, mstot, depressive symptoms and sample weight
```{r model_selection1, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 3
```{r model_selection2, echo=C_echo, message=C_message, warning=C_warning}
```

##Model Selection
```{r model_selection3, echo=C_echo, message=C_message, warning=C_warning}
```

##Model Conclusion
Model Selection gave ambiguous responsed. From visual examination, it was determined that a model of ~6 variables would be appropriate. These variables were education, age, cognition, number of health conditions,  race and self-rated health score

##Slide 4
```{r add_graphs1, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 5
```{r add_graphs2, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 6
```{r add_graphs3, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 7
```{r add_graphs4, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 8
```{r add_graphs5, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 9
```{r add_graphs6, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 10
```{r add_graphs7, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 11
```{r add_graphs8, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 12
```{r add_graphs9, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 13
```{r add_graphs10, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 14
```{r add_graphs11, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 15
```{r add_graphs12, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 16
```{r add_graphs13, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 17
```{r add_graphs14, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 18
```{r add_graphs15, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 19
```{r add_graphs16, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 20
```{r add_graphs17, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 21
```{r add_graphs18, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 22
```{r add_graphs19, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 23
```{r add_graphs20, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 24
```{r add_graphs21, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 25
```{r add_graphs22, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 26
```{r add_graphs23, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 27
```{r add_graphs24, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 28
Creates models of variables of interest
```{r define_variables1, echo=TRUE, message=C_message, warning=C_warning}
```

##Slide 29
```{r define_variables2, echo=TRUE, message=C_message, warning=C_warning}
```

##Slide 30
```{r define_variables3, echo=TRUE, message=C_message, warning=C_warning}
```

##Slide 31
```{r define_variables4, echo=TRUE, message=C_message, warning=C_warning}
```

##Slide 32
Adds models to figures produced above
```{r add_model_data1, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 33
```{r add_model_data2, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 34
```{r add_model_data3, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 35
```{r add_model_data4, echo=C_echo, message=C_message, warning=C_warning}
```



##Slide 36
Health conditions
Data show that the more health conditions you have, the higher your BMI. Given the high co-incidence of obesity (see Patterson et al., 2004) with certain health conditions such as diabetes and heart disease, this is not a surprising finding.

Age
Data show that the older you are, the lower your BMI is. This is not a suprising finding, given that people who live longer tend to have better health, which would ideally correlate with a lower BMI.

Cogtot
Data show that the higher your cognition, the lower your BMI.

Education
Data seem show that the more years of education you have, the higher your BMI. This is somewhat suprising given that conventional wisdom (my bias?) says that education tends to correlate with wealth which correlates with better health. This could either indicate a problem with my conflation of BMI as a measurement of overall health or with differences in the retired population.

Race
Race seems to show that being african american correlates with a higher BMI, this is not a super suprising finding, somewhat expected.

Drink
Drinking data seems to show that when participants drink, they are more likely to have a lower BMI. This is a very suprising finding, although it could indicate a problem with their definition of drinking. Unfortunately I was unable to find the operation definition of if a participant drank or not, which could mean either quite a bit of drinking or just a certain number of drinks per week. 

Shlt
Participants rating of their health (self-rated health score) indicates that people who rate their health lower tend to have a higher BMI.
