---
title: "Final Project: PSYC 532"
author: "Pauline Song"
date: "December 7, 2015"
output:
  ioslides_presentation:
    smaller: yes
---  

```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_analysis/data_analysis_hrs.R") # the file to which knitr calls for the chunks
```


##Graphical Exploration of Multiple Variables Predicting One Outcome


```{r LoadPackages, echo=F, message=C_message, warning=C_warning}
```


```{r LoadData, echo=F, message=C_message, warning=C_warning}
```


```{r LoadSources, echo=F, message=C_message, warning=C_warning}
```

##Selecting Variables
Select desired variables and verify the pulled variables. 
```{r inspectData, echo=C_echo, message=C_message, warning=C_warning}
```

##Class Change
Add new variables to change the the class of the variables to factor/categorical.
```{r inspectData2, echo=C_echo, message=C_message, warning=C_warning}
```

##Model
Created a model to encompass all the factors of the dataset: BMI,  All the factors in the model are significant predictors of self reported health
```{r BasicModel, echo=C_echo, message=C_message, warning=C_warning}
```

##Graphical Exploration


##Perceived Health Deterioration & Psychiatric Condition
```{r BasicGraph1, echo=F, message=C_message, warning=C_warning}
```

##Perceived Health Deterioration & Lung Condition
```{r BasicGraph3, echo=F, message=C_message, warning=C_warning}
```

##Perceived Health Deterioration & Heart Condition
```{r BasicGraph4, echo=F, message=C_message, warning=C_warning}
```

##Perceived Health Deterioration & Cancer
```{r BasicGraph5, echo=F, message=C_message, warning=C_warning}
```

##Perceived Health Deterioration & Diabetes
```{r BasicGraph6, echo=F, message=C_message, warning=C_warning}
```

##Exploration of Multiple Variables Contributing to Self-Reported Health

##Basic Graph: Perceived Health Deterioration vs. BMI
```{r BasicGraph7, echo=F, message=C_message, warning=C_warning}
```

##Heart, Lung, Diabetes & BMI as Predictors for Health Deterioration
```{r BasicGraph8, echo=F, message=C_message, warning=C_warning}
```

##Psych, Heart, Lung & BMI as Predictors for Health Deterioration
```{r BasicGraph9, echo=F, message=C_message, warning=C_warning}
```

##Diabetes, Lung, Heart & BMI as Predictors for Health Deterioration
```{r BasicGraph10, echo=F, message=C_message, warning=C_warning}
```

##Psych, Heart, Lung & BMI as Predictors for Health Deterioration
```{r BasicGraph11, echo=F, message=C_message, warning=C_warning}
```

##Diabetes, Heart, Psych & BMI as Predictors for Health Deterioration
```{r BasicGraph12, echo=F, message=C_message, warning=C_warning}
```

##Lung, Heart, Psych & BMI as Predictors for Health Deterioration
```{r BasicGraph13, echo=F, message=C_message, warning=C_warning}
```

##Any tentative conclusions? Further explorations?
--> Heart condition OR Lung condition?

##Thank you!
```{r SaveDerivedData, echo=C_echo, message=C_message, warning=C_warning}
```

```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```
