---
title: "Final Project"
author: "Thomas Ferguson"
date: "December 5, 2015"
output: 
  ioslides_presentation: 
    smaller: yes
    widescreen: yes
---

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- FALSE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_analysis/basic_analysis.R") # the file to which knitr calls for the chunks
```

##Load Packages
Prepare RStudio environment for all tasks to follow. 
```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```

```{r load_sources, echo=C_echo, message=C_message, warning=C_warning}
```

```{r DeclareGlobals, echo=C_echo, message=C_message, warning=C_warning}
```

We load the data derived by the script "./reports/data_preparation/dsL_hrs.R"" 
Variables chosen are age in years, eduction in years, gender, race, if they drink currently, if they smoked ever,
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 1
Load graph settings for creating figures
```{r LoadGraphSettings, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 2
Chooses variables for model - End up with model of ~6 or 7 due to conflicting results (seemed reasonable)
```{r model_selection, echo=C_echo message=C_message, warning=C_warning}
```

##Slide 3
Create Plots of cognition, number of chronic conditions, education in years and age in years by BMI (with gender, if they drank, and race included as factors)
```{r add_graphs, echo=C_echo, message=C_message, warning=C_warning}
```


##Slide 4
Creates prediction function for models
```{r get_prediction_function, echo=C_echo, message=C_message, warning=C_warning}
```


##Slide 5
Creates models of variables of interest
```{r define_models, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 6
Adds models to figures produced above
```{r add_model_data, echo=C_echo, message=C_message, warning=C_warning}
```

##Slide 7 - Conclusions

## Appendix

Removes all variables except variables conisdered for models
```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```

