---
title: "SATSA Data Preparation"
author: "Jamie Knight"
date: "November 26, 2015"
output: html_document
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: SATSA
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_preparation/data_preparationFINAL.R") # the file to which knitr calls for the chunks
```

##Introduction

In this report we are preparing a derived data set using the SATSA data available online at (http://www.icpsr.umich.edu/icpsrweb/DSDR/studies/3843). 

Within the SATSA study a subsample of the participants were interviewed in seven waves of in-person testing and those measurements included smell and taste questions along with cognitive and functional capacity.

This subsample is what we will use in our analysis and includes:  
- Questionnaire 3 and In-Person Testing Time 3 (1990);   
- In-Person Testing Time 5 (2004);   
- Questionnaire 6 and In-Person Testing Time 6 (2007);   
- In-Person Testing Time 7 (2010)  

In this script we will derive a data set which will include smell, taste, sex, cognitive MMSE scores and id variables. I will also explore an issue with the original coding of the data that could potentially have consequences for data analysis. 


##Load Packages
This section prepares the RStudio environment for all tasks to follow.

If the packages are not insalled:  
(1) Uncomment the install.packages list by highlighting them and holding down Shift CTRL C  
(2) Run the code   
(3) Comment out the insall.packages list using the same process as #1.   
(4) Load the packages.  

Installing packages will only need to be done once.
Loading packages will need to be done at the beginning of each session.
```{r LoadPackages, message=C_message, warning=C_warning}
```

##Raw Data 
If you haven't done so already, you will need to sign up and download the required raw data from the SATSA website (http://www.icpsr.umich.edu/icpsrweb/DSDR/studies/3843). Directions for this process can be found in the README file in the raw data folder. 

##Load and Rename Data
We are going to load, rename and subset the data to just the variables we need: smell and taste across waves 3, 5, 6, and 7.

In this section we begin to see that there is a major issue with the coding for the questionnaires across waves.
```{r LoadRename, echo=TRUE, message=C_message, warning=C_warning}
```

##Frequency Tables 
The frequncy tables for smell for each wave are shown here.
You can clearly see how they have coded each wave differently using a combination of numbers and words that do not match between waves.

This means that smell is not an ordered variable across waves even though it is within each wave. This makes longitudinal analysis very difficult.
```{r show_values, echo=TRUE, message=C_message, warning=C_warning}
```

##Recode Sex
```{r RenameSex, echo=TRUE, message=C_message, warning=C_warning}
```

##Recode Smell and Taste
As you saw above the coding for these is not ordinal.  
Each wave has a different question they are asking when it comes to smell and taste.  
In one wave they ask: "How is your smell?" with the responses being either Good, Pretty Good or Bad.

In the next wave they wil ask the same question but the respons options being: Good, Not So Good, or Bad.

This means that we cannot use smell and taste across waves, as the answers are not equivalent.

##Recode Anyway
We are going to recode them as if they are ordinal with Good, Pretty Good, Not So Good, and Bad being equivalent to a score of 1, 2, 3 and 4 respectively in order to see what the individual growth plots look like over the waves for each individual. I would normally expect to see a linear decline. We can see these growth plots in the next section called data_analysis. 

```{r RenameVariables, echo=TRUE}
```
##Save Derived Data
```{r SaveDerivedData, echo=TRUE}
```
##Clean Up 
```{r CleanUp, echo=TRUE}
```

