---
title: "PSYC 532 Data Preparation Report"
output:
  html_document:
    toc: yes
    self_containted: false
    keep_md: true
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: PSYC532 Final Project
BRANCH: the-name-of-the-branch
PATH: ../reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_preparation/hrs_final/hrs_final.R") # the file to which knitr calls for the chunks
```
##Set Up Environment
###Load Packages
Prepare RStudio environment for all tasks
```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```

###Load Data
Link to the data source
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

###Load Graph Settings
```{r LoadGraphSettings, echo=C_echo, message=C_message, warning=C_warning}
```

##Modify Data
###Rename Variables
Rename variables for easier handling
```{r RenameVariables, echo=C_echo, message=C_message, warning=C_warning }

```

###Group by ID
```{r GroupID, echo=C_echo, message=C_message, warning=C_warning }
```

###Sample Data
```{r SubsetData, echo=C_echo, message=C_message, warning=C_warning}
```

###Graph Outcome Variable
We see relatively even distribution for BMI
```{r BasicGraph, echo=C_echo, message=C_message, warning=C_warning}
```

###Filter Variables
Select variables that are necessary for analysis: id, wave, age, sex, bmi, arthr
```{r DataForModel, echo=C_echo, message=C_message, warning=C_warning}
```

###Remove Missing Data
Remove missing data so that only complete cases are left
```{r RemoveMissing, echo=C_echo, message=C_message, warning=C_warning}
```

##Finishing Touches
###Save Derived Data
Save new dataset for future analysis
```{r SaveDerivedData, echo=C_echo, message=C_message, warning=C_warning}
```

###Remove Other Datasets
Keep only the final dataset
```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```