---
title: "Data Analysis"
output:
  html_document:
    toc: yes
    self_containted: false
    keep_md: true
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: the-name-of-the-repository
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_analysis/data_analysis.R") # the file to which knitr calls for the chunks
```

#Load R Packages
```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```

#Load graphing sources
```{r LoadSources, echo=C_echo, message=C_message, warning=C_warning}
```

#Load tidy dataset
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

#Exploring the data
##Histograms
```{r histogram, echo=C_echo, message=C_message, warning=C_warning}
```

##Scatterplots
To view relationship between BMI and other continuous variables

```{r Scatterplot1, echo=C_echo, message=C_message, warning=C_warning}
```

The general trend of this scatterplot suggests that cognition increases as BMI increases

```{r Scatterplot2, echo=C_echo, message=C_message, warning=C_warning}
```

The general trend of this scatterplot suggests that as individuals get older, their BMI decreases

```{r Scatterplot3, echo=C_echo, message=C_message, warning=C_warning}
```

The general trend of this scatterplot suggests that as years of education increases, BMI decreases 


To view a snapshot of the data.
```{r Dataview, echo=C_echo, message=C_message, warning=C_warning}
```

#Creating a Model to Predict BMI
##Best subset selection
Regsubsets function performs best-subset selection by identifying the best model that contains a given number of predictors, where best is defined using RSS.
Best subset selection fits a separate least squares regression for each possible combination of the p predictors.
```{r Bestsubsetselection, echo=C_echo, message=C_message, warning=C_warning}
```

##Best subset selection - forward stepwise
Regsubsets function can also perform forward stepwise selection by using the method="forward". Forward stepwise begins with a model containing no predictors, and then adds predictors to the model one-at-a-time, until all of the predictors are in the model.
```{r SubsetSelectionForward, echo=C_echo, message=C_message, warning=C_warning}
```

##Best subset selection - backward stepwise
Use function regsubsets and method="backward". Backward stepwise begins with the full least squares model containing all p predictors, and then removes the least useful predictor, one-at-a-time.
```{r SubsetSelectionBackward, echo=C_echo, message=C_message, warning=C_warning}
```

All three methods of best subset selection produce same results.   

##Using Adjusted R-Squared
A large value of adjusted r-squared indicates a model with a small test error.
The which.max() function can be used to identify the location of the maximum point of a vector.
Adjusted r-squared says that a model with 5 factors will produce the largest adjusted r-squared.

```{r adjustedrsquared, echo=C_echo, message=C_message, warning=C_warning}
```

##Using Cp
Cp is an unbiased estimate of test MSE. The Cp statistic takes on a small value for models with low test error.
The which.min() function can be used to identify the location of the minimum point of a vector.
Cp says that a model with 4 factors will produce the smallest Cp.
```{r cp, echo=C_echo, message=C_message, warning=C_warning}
```

##Using BIC
Bayesian information criterion (BIC) will take on a small value for a model with a
low test error, so we select the model that has the lowest BIC value.
The which.min() function can be used again to idenify the location of the minimum point of the vector.
BIC says that a model with 4 factors suggests that 4 factors optimal.
```{r bic, echo=C_echo, message=C_message, warning=C_warning}
```


###Plots of all 3 methods
Plot Adjusted r-squared with a point on the which.max() value calculated above
```{r plottingadjr2, echo=C_echo, message=C_message, warning=C_warning}
```

Plot Cp with a point on the which.min() value calculated above
```{r plottingcp, echo=C_echo, message=C_message, warning=C_warning}
```

Plot BIC with a point on the which.min() value calculated above
```{r plottingbic, echo=C_echo, message=C_message, warning=C_warning}
```

#Investigating which model best predicts BMI
#Linear model with all 6 factors

```{r Analysis1, echo=C_echo, message=C_message, warning=C_warning}
```

The linear model with all 6 factors shows that all variables are significant predictors of BMI except for psychiatric issues and cognition.

#Linear model with 5 factors
Relationship of BMI with vigorous activity, age, gender, years of education, and psychiatric issues.
These variables were chosen by looking at best subset selection.
This is the optimal model size suggested by adjusted r-squared which.max() function.
```{r Analysis2, echo=C_echo, message=C_message, warning=C_warning}
```
The linear model with 5 variables shows that all variables are significant predictors of BMI at the .001 level except for psychiatric issues.


#Linear model with 4 factors
Relationship of BMI with vigorous activity, age, gender, and years of education.
Variables were chosen by looking at best subset selection.
This is the optimal model size suggested by Cp and BIC which.min() function
```{r Analysis3, echo=C_echo, message=C_message, warning=C_warning}
```
The linear model with 4 variables shows that all variables are significant predictors of BMI

#Linear model with 3 factors
Relationship of BMI with age, years of education and vigorous activity
Variables were chosen by looking at best subset selection.
```{r Analysis4, echo=C_echo, message=C_message, warning=C_warning}
```

#Linear model with 2 factors
Relationship of BMI with age and years of education
Variables were chosen by looking at best subset selection.
```{r Analysis5, echo=C_echo, message=C_message, warning=C_warning}
```

#Linear model with 1 factor
Relationship of BMI with age. Age was chosen as the factor to use in the one variable model after looking at best subset selection.
```{r Analysis6, echo=C_echo, message=C_message, warning=C_warning}
```

#Linear model with cognition
Wanted to explore whether cognition would be a significant predictor on its own in a model since it was not significant in the model with 6 factors
```{r Analysis7, echo=C_echo, message=C_message, warning=C_warning}
```
The results show that cognition would be considered a significant predictor of BMI if it was the only factor in a model with BMI

#Linear model with psychiatric issues
Wanted to explore whether psychiatric issues would be a significant predictor on its own in a model since it was not significant in the model with 5 or 6 factors
```{r Analysis8, echo=C_echo, message=C_message, warning=C_warning}
```

#Linear model with cognition and psychiatric issues
Wanted to explore whether both cognition and psychiatric issues would be significant predictors of BMI on their own in a model since they weren't significant in the model with 6 factors
```{r Analysis9, echo=C_echo, message=C_message, warning=C_warning}
```
Results show that cognition and psychiatric issues would be considered significant predictors of BMI if in a model with just those 2 factors

#Final linear model
The final model I chose to predict BMI is the model with 4 factors (age, years of education, gender, and vigorous activity). Since both BIC and Cp had the lowest values with this model size, and best subset selection (all 3 methods) suggested that these 4 factors should be the factors included in this size model, and all factors within this model are statistically significant predictors of BMI, I think this is the most simple and most accurate model to use to best predict BMI.

```{r Finalanalysis, echo=C_echo, message=C_message, warning=C_warning}
```

The coefficients tell us the as BMI increases by 1 point, vigorous activity decreases by (.82, SD=.103), age decreases by (.15, SD=.007), the effect of being male increases by (.63, SD=.101), and years of education decreases by (.14, SD=.015).

#Conclusions

- From the 6 initial predictors in my dataset, I found that the model with age, years of education, vigorous activity and gender best predicts BMI

- The coefficients tell us that as BMI increases:
- age decreases (what our initial scatterplot showed)
- years of education decreases (what our initial scatterplot showed)
- effect of being male increases
- vigorous activity decreases

- I also discovered how misleading linear models can be when I looked into cognition and psychiatric issues on their own

- Although cognition and psychiatric issues appear to be significant predictors of BMI when in a model on their own, when looking at the more complex model with 6 predictors, they are no longer significant predictors of BMI

