---
title: "Data Preparation"
output:
  html_document:
    toc: yes
    self_containted: false
    keep_md: true
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: the-name-of-the-repository
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_preparation/data_preparation.R") # the file to which knitr calls for the chunks
```

```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```
#Load [Healthy Retirement Study Data](http://hrsonline.isr.umich.edu/)
```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```
#Load themes used to style graphs
```{r LoadGraphSettings, echo=C_echo, message=C_message, warning=C_warning}
```
#Select variables from dataset
Selected variables that look at psychiatric issues, gender, vigorous activity, cognition, age, years of education, and bmi. Also included 'wave' variable so I could pick one timepoint in the dataset to look at

[Information on assessing BMI](http://www.cdc.gov/healthyweight/assessing/bmi/)

I am interested in looking at which variables are predictive of BMI. 
```{r selectdsM, echo=C_echo, message=C_message, warning=C_warning}
```

#Look at summary of variables
```{r summary, echo=C_echo, message=C_message, warning=C_warning}
```

##Delete missing information
```{r deletemissinginfo, echo=C_echo, message=C_message, warning=C_warning}
```

##Select one timepoint (wave=4)
```{r Timepoint, echo=C_echo, message=C_message, warning=C_warning}
```

#Look at summary of variables now that missing info is gone and one timepoint has been selected
Gender, psychiatric issues, and vigorous activity are all categorical variables
```{r newsummary, echo=C_echo, message=C_message, warning=C_warning}
```

#Save derived data
```{r SaveDerivedData, echo=C_echo, message=C_message, warning=C_warning}
```
#Remove all datasets except for dsM
```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```
