---
title: "Analyzing the Community Health Status Indicator Data"
output:
  html_document:
    toc: yes
    theme: spacelab
    font-family: helvetica
    self_containted: false
    keep_md: true
    runtime: shiny
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: the-name-of-the-repository
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./scripts/data/shiny_script.R") # the file to which knitr calls for the chunks
```

```{r LoadPackages, echo=C_echo, message=C_message, warning=C_warning}
```

```{r LoadData, echo=C_echo, message=C_message, warning=C_warning}
```

```{r, DataPrepForShiny, echo=FALSE, results='hide'}
```

####First, select the variable you want to investigate.
```{r, ShinyUI, echo=FALSE}
```



####And here's the histogram for that variable.
```{r, ShinyServer, echo=FALSE}
```

***
***
***
***
***
***
***
***
***
***
***




####Using **ggvis**, you can create more detailed and complex interactive plots. Here's a histogram with adjustable bins!
```{r, ggvis, echo=FALSE}
```


***
***
***
***
***
***
***
***
***
***
***



####Here's a plot of _Obesity by number people reporting high blood pressure (per 100,000)_. The span of the smoothing curve can be adjusted, as well as the size and colour of the points. 
```{r, ggvis2, echo=FALSE}
```
