---
title: "Data preperation for final report - Thomas Ferguson"
output:
  html_document:
    toc: yes
    self_containted: false
    keep_md: true
---  
<!-- for more options study http://rmarkdown.rstudio.com/html_document_format.html  -->
<!-- The report is produced from
REPOSITORY: Finalproject
BRANCH: the-name-of-the-branch
PATH: ../Reports/
-->

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of only one directory.-->
```{r, echo=F, message=F} 
library(knitr)
opts_knit$set(root.dir='../../')  #Don't combine this call with any other chunk -especially one that uses file paths.
```

<!-- Set the report-wide options, and point to the external script file. -->
```{r, echo=F, message=T}
require(knitr)
opts_chunk$set(
  results='show', # Set
  message = TRUE, # De
  comment = NA,   #   fau
  tidy = FALSE,   #     lts
#   fig.height = 4.8, # physical dimensions
#   fig.width = 6.5,  # of produced impage
  out.width = NULL, # how it looks in html document
  fig.path = 'figure_rmd/', # where images from this report are stored
  dev = "png",
  dpi = 100 # resolution: dots per inch. Use 100 to develop, 300-400 to print. adjust
)
C_echo <- TRUE
C_message<- FALSE
C_warning<- FALSE

options(width=120) #So the output is 50% wider than the default.
read_chunk("./reports/data_preperation/dsL_hrs.R") # the file to which knitr calls for the chunks
```

To load desired packages
```{r load_packages, echo=C_echo, message=C_message, warning=C_warning}
```

```{r load_sources, echo=C_echo, message=C_message, warning=C_warning}
```

Find file path for data
```{r declare_global, echo=C_echo, message=C_message, warning=C_warning}
```

Set file path for data
```{r load_data, echo=C_echo, message=C_message, warning=C_warning}
```

Examine data and select wave (describes what portion of sample...) - done because of file size
```{r inspect_data, echo=C_echo, message=C_message, warning=C_warning}
```

```{r tweak_data, echo=C_echo, message=C_message, warning=C_warning}
```

Inspect data and examine relationships between data
```{r basic_table, echo=C_echo, message=C_message, warning=C_warning}
```

```{r query_data, echo=C_echo, message=C_message, warning=C_warning}
```

Create vector  containing only variables of interest to BMI
#These variables are 
```{r arrivedsW1, echo=C_echo, message=C_message, warning=C_warning}
```

```{r remove_na, echo=C_echo, message=C_message, warning=C_warning}
```

```{r mutate_data1, echo=C_echo, message=C_message, warning=C_warning}
```

```{r mutate_data2, echo=C_echo, message=C_message, warning=C_warning}
```

```{r mutate_data3, echo=C_echo, message=C_message, warning=C_warning}
```

```{r mutate_data4, echo=C_echo, message=C_message, warning=C_warning}
```

```{r mutate_data5, echo=C_echo, message=C_message, warning=C_warning}
```

```{r arrivedsW2, echo=C_echo, message=C_message, warning=C_warning}
```

Load graph settings for creating figures
```{r LoadGraphSettings, echo=C_echo, message=C_message, warning=C_warning}
```

Use GAM or Stepwise selection to select best model - Variables chosen are age in years, number of health conditions, eduction in years and self-rated health score
```{r model_selection, echo=C_echo, message=C_message, warning=C_warning}
```

Create figures for modeling
```{r add_models_smoothers, echo=C_echo, message=C_message, warning=C_warning}
```

Creates prediction function for models
```{r get_prediction_function, echo=C_echo, message=C_message, warning=C_warning}
```

Creates models of variables of interest
```{r define_models, echo=C_echo, message=C_message, warning=C_warning}
```

Adds models to figures produced above
```{r add_model_data, echo=C_echo, message=C_message, warning=C_warning}
```

## Conclusions


### Health conditions

Data show that the more health conditions you have, the higher your BMI. Given the high co-incidence of obesity (see Patterson et al., 2004) with certain health conditions such as diabetes and heart disease, this is not a surprising finding.

### Age

Data show that the older you are, the lower your BMI is. This is not a suprising finding, given that people who live longer tend to have better health, which would ideally correlate with a lower BMI.

###Cogtot

Data show that the higher your cognition, the lower your BMI.

###Education

Data seem show that the more years of education you have, the higher your BMI. This is somewhat suprising given that conventional wisdom (my bias?) says that education tends to correlate with wealth which correlates with better health. This could either indicate a problem with my conflation of BMI as a measurement of overall health or with differences in the retired population.

###Race

Race seems to show that being african american correlates with a higher BMI, this is not a super suprising finding, somewhat expected.

###Drink

Drinking data seems to show that when participants drink, they are more likely to have a lower BMI. This is a very suprising finding, although it could indicate a problem with their definition of drinking. Unfortunately I was unable to find the operation definition of if a participant drank or not, which could mean either quite a bit of drinking or just a certain number of drinks per week. 


###Shlt

Participants rating of their health (self-rated health score) indicates that people who rate their health lower tend to have a higher BMI.

## Appendix

Removes all variables except variables conisdered for models
```{r arrivedsW3, echo=C_echo, message=C_message, warning=C_warning}
```

Creates an rds and a csv file that contains only the variables of interest for models (reduces data set size)
```{r SaveDerivedData, echo=C_echo, message=C_message, warning=C_warning}
```

```{r CleanUp, echo=C_echo, message=C_message, warning=C_warning}
```